QPLIB: A Library of Quadratic Programming Instances

### 2018-08-23 AnswerToTheReferees2_second.tex

\rep{We would like to thank the AE for his/her positive opinion on the revision we performed of our manuscript. 
%\leo{Togliamo la spiegazione degli algoritmi e sostituiamola con una citazione di un book-tutorial-survey-introduction.}
}

\bigskip
\textbf{Main points:}

\begin{itemize}
\item 
{\it
[from referee, paraphrased] 8. Please try to find a more precise definition of asypmtotically complete algorithm to 
cite.
}

\rep{We have added the definition in paragraph 2.3.3 and it is taken from: 
\begin{itemize}
\item Neumaier, A.: Complete search in continuous global optimization and
constraint satisfaction. Acta Numerica 13, 271–369 (2004)
\end{itemize}
}

### 2018-08-24 QPLIB-3.tex
proofreading, mainly Section 3

\textcolor{green}{We chose instances based on the following four goals:}
%During the development of the library, a discussion ensued about
%the expected goals that we wished to achieve. The following four goals
%were finally identified:
%
\begin{enumerate}
 \item to represent as far as possible all the different categories of QP
       problems;
 %
 \item to gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item to include, for each of the categories, a set of well-diversified
       instances; and
 %
 \item to obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve such goals, we performed the following two filters, applied in
a cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S\ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instances we
       ran the complete solvers in GAMS, whose number depends on the category
       of the instance under consideration, cf.~Table \ref{t:solvers}.
       \textcolor{green}{Thanks to these extensive preliminary tests, we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.}
       %We then filtered according to a first
       %measure of computational difficulty, i.e., we discarded all instances
       %that are solved by at least 30\% of the complete solvers within a time
       %limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       \textcolor{green}{We carefully analyzed the instances one by one,
       %, and we clustered them
       %according to their features; for each cluster we kept only a few
       %representatives, e.g. 
        eliminating all but a few of those with very similar size and coming from the same donor.
       The instances discarded by this second filter are instances of the same specific problem, e.g., we gathered many Max Cut Problem instances and we kept in the library only a representative small set of them. The selected representative instances are the larger and computationally harder ones.}
       Finally, in order to only
       keep computationally challenging instances we ran a complete solver for \textit{QGQ} with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded.
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $319$ discrete instances and
$134$ continuous instances.

\begin{table}
 \centering
 \setlength{\tabcolsep}{5pt}
%  \arraystretch{1}
\begin{tabular}{cccc}
Starting set& \multicolumn{ 2}{c}{ $\approx$ 8500 Instances }& \\
& \multicolumn{ 2}{c}{$\Downarrow$}& \\
& $\approx$ 6000 Discr. Inst.  & $\approx$ 2500 Cont. Inst. & \\
First Filter  & $\Downarrow$  & $\Downarrow$ & \\
 & $\approx$ 3000 Discr. Inst.  & $\approx$ 1000 Cont. Inst. & \\
Second Filter & $\Downarrow$  & $\Downarrow$  & \\
% & 600 Discr. Inst.  & 250  Cont. inst. & \\
  & 319 Discr. Inst.  & 134  Cont. inst. & \\
\end{tabular}
%\begin{center}\end{center}
\caption{Instance filter steps} \label{tab:filters}
\end{table}

@@

{We now analyze the features of the instances selected to be part of the
 library. Table \ref{tab:DDDDD} provides a global overview. The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories. As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise. On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex (if minimization) or nonconcave (if maximization) and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise. }

@@

 The detailed characteristics of the instances are presented in Table
\ref{tab:DD} for \emph{discrete} instances
(\textit{*}\{\textit{B,M,I,G}\}\textit{*}) and in Table
\ref{tab:CC} for \emph{continuous} ones (\textit{*C*}).
For each category, the tables
report in column ``$\#$'' the corresponding number of instances. It can be seen
that the final set well respects the original distribution of the gathered
instances among the different categories. Indeed, the discrete categories
\textit{LMQ} and \textit{QBL} are well represented by {134} and {91}
instances, respectively. Similarly, the continuous categories
\textit{LCQ} and \textit{QCQ} are well
represented by {52} and {30} instances, respectively. Moreover, the library
actually covers the large majority of all possible categories of instances.

@@

We now report some graphs that help in illustrating the main features
of the instances. In Figure~\ref{fig:distribution}~(left) we plot the number
of variables (horizontal axis) versus the number of constraints
(vertical axis), both in logarithmic scale. Continuous instances are
marked with ``$+$'', and discrete ones with ``$\times$''. The figure shows
that the library contains a quite diverse set of instances in
terms of number of variables and constraints. {The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances QPLIB\_8547 and QPLIB\_9008.}
Figure~\ref{fig:distribution}~(right) plots the number of nonzero elements in the gradient of the objective function and the Jacobian and the number of these nonzeros corresponding to nonlinear variables, that is, it counts the appearances of variables in objectives and constraints and how often such an appearance is in a quadratic term.

>>>>

\textcolor{green}{We chose instances based on the following four goals:}
%During the development of the library, a discussion ensued about
%the expected goals that we wished to achieve. The following four goals
%were finally identified:
%
\begin{enumerate}
 \item to represent as far as possible all the different categories of QP
       problems;
 %
 \item to gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item to include, for each of the categories, a set of well-diversified
       instances; and
 %
 \item to obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve such goals, we performed the following two filters, applied in
a cascade:
%
\begin{itemize}
 \item \emph{First Instance Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S\ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instances we
       ran the complete solvers in GAMS, whose number depends on the category
       of the instance under consideration, cf.~Table \ref{t:solvers}.
       \textcolor{green}{Thanks to these extensive preliminary tests, we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.}
       %We then filtered according to a first
       %measure of computational difficulty, i.e., we discarded all instances
       %that are solved by at least 30\% of the complete solvers within a time
       %limit of 30 seconds.
 %
 \item \emph{Second Instance Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       \textcolor{green}{We carefully analyzed the instances one by one,
       %, and we clustered them
       %according to their features; for each cluster we kept only a few
       %representatives, e.g. 
        eliminating all but a few of those with very similar size and coming from the same donor.
       The instances discarded by this second filter are instances of the same specific problem, e.g., we gathered many Max Cut Problem instances and we kept in the library only a representative small set of them. The selected representative instances are the larger and computationally harder ones.}
       Finally, in order to only
       keep computationally challenging instances we ran a complete solver for \textit{QGQ} with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded.
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $319$ discrete instances and
$134$ continuous instances.

\begin{table}
 \centering
 \setlength{\tabcolsep}{5pt}
%  \arraystretch{1}
\begin{tabular}{cccc}
Starting set& \multicolumn{ 2}{c}{ $\approx$ 8500 instances }& \\
& \multicolumn{ 2}{c}{\rotatebox{-45}{$\Downarrow$} \quad \raisebox{-1.2ex}{\rotatebox{45}{$\Downarrow$}}}& \\
& $\approx$ 6000 discr. inst.  & $\approx$ 2500 cont. inst. & \\
First filter  & $\Downarrow$  & $\Downarrow$ & \\
 & $\approx$ 3000 discr. inst.  & $\approx$ 1000 cont. inst. & \\
Second filter & $\Downarrow$  & $\Downarrow$  & \\
% & 600 Discr. Inst.  & 250  Cont. inst. & \\
  & 319 discr. inst.  & 134  cont. inst. & \\
\end{tabular}
%\begin{center}\end{center}
\caption{Instance filter steps} \label{tab:filters}
\end{table}

@@

{We now analyze the features of the instances selected to be part of the
 library. Table \ref{tab:DDDDD} provides a global overview. The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories. As previously mentioned, an instance is classified \textit{discrete} if it contains at least one binary or integer variable, and \textit{continuous} otherwise. On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex (if minimization) or nonconcave (if maximization) and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise. }

@@

 The detailed characteristics of the instances are presented in Table
\ref{tab:DD} for \emph{discrete} instances
(\textit{*}\{\textit{B,M,I,G}\}\textit{*}) and in Table
\ref{tab:CC} for \emph{continuous} ones (\textit{*C*}).
For each category, the tables
report the corresponding number of instances in column ``$\#$''. It can be seen
that the final set well respects the original distribution of the gathered
instances among the different categories. Indeed, the discrete categories
\textit{LMQ} and \textit{QBL} are well represented by {134} and {91}
instances, respectively. Similarly, the continuous categories
\textit{LCQ} and \textit{QCQ} are well
represented by {52} and {30} instances, respectively. Moreover, the library
actually covers the large majority of all possible categories of instances.

@@

We now report some graphs that help in illustrating the main features
of the instances. In Figure~\ref{fig:distribution}~(left) we plot the number
of variables (horizontal axis) versus the number of constraints
(vertical axis), both in logarithmic scale. Continuous instances are
marked with ``$+$'' and discrete ones with ``$\times$''.
\textcolor{green}{Box constraints are not counted as constraints.} \todo{Ambros: From text, tables, and Figure~1 did I correctly infer that variable bounds are not counted as constraints?  The 23 ``none'' in the discrete table seem to be the blue crosses on the x-axis, but then where do the 3 red ``+'' on the x-axis come from?  Are these continuous instances with box constraints and some boxes are infinite?  Should there be a ``none'' in the continuous table?}
The figure shows
that the library contains a quite diverse set of instances in
terms of number of variables and constraints. {The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances QPLIB\_8547 and QPLIB\_9008.}
Figure~\ref{fig:distribution}~(right) plots the number of nonzero elements in the gradient of the objective function and the Jacobian and the number of these nonzeros corresponding to nonlinear variables, that is, it counts the appearances of variables in objectives and constraints and how often such an appearance is in a quadratic term.
